Quantum Hoare Type Theory

### 2020-10-25 ms-paper/ms-paper.tex
Metavariable M is now used for operations as in M.P, could use U perhaps?

\begin{center}
	\begin{tabular}{ll}
		$\Delta \vdash K \syn A [N']$ & Elim term K has type A and canonical form $N'$ \\
		$\Delta \vdash M \chk A [M']$ & Intro term M has type A and canonical form $M'$
	\end{tabular}
\end{center}

@@

\paragraph{Initialization}
$x \Leftarrow \kw{mkQbit} M; E$

\begin{mathpar}
	\inferrule
	{\Delta \vdash M \chk \type{Bool}
		\\ \Delta, x : \type{Qbit}; P \circ (x \mapsto \kw{state}(M)) \vdash E \syn y:B . Q}
	{\Delta; P \vdash x \Leftarrow \kw{mkQbit} M; E \syn y:B. (\exists x:\type{Qbit}.Q)}
\end{mathpar}
where x is fresh and \texttt{state} is a function that translates a classical representation of quantum state (such as boolean here) to the equivalent quantum state vector.

Initialization can only be performed for a term $M$ that can be type checked as a \type{Bool}. Then, we look at the rest of the computation, $E$, in a context that adds $x$ of type \type{Qbit} under a precondition that extends the previous precondition, $P$, with the strongest postcondition for initialization, that is, the newly bound variable $x$ points to the representation of a new qubit state; this should synthesize the strongest postcondition (with respect to the expanded precondition in the context), $Q$, for $E$.

@@

\paragraph{Measurement}
$x \Leftarrow \kw{measQbit} M; E$

\begin{mathpar}
	\inferrule
	{\Delta \vdash M \chk \type{Qbit}
		\\ \Delta; \Psi; P \prove (M \hookrightarrow -)
		\\ \Delta, x:\type{Bool}; P \circ ((M \mapsto -) \multimap \kw{emp}) \vdash E \syn y:B . Q}
	{\Delta; P \vdash x \Leftarrow \kw{measQbit} M; E \syn y:B. (\exists x:\type{Bool}.Q)}
\end{mathpar}

The measurement rule can be read in a similar way as the previous rule. But in measurement, we need to additionally prove that the verification condition (written as a sequent), the precondition $P$ implies that the qubit $M$ is allocated, holds in the heap context $\Psi$. The strongest postcondition here is that the fragment of heap that only refers to location M becomes empty.

>>>>

\begin{center}
	\begin{tabular}{ll}
		$\Delta \vdash K \syn A [N']$ & Elim term K has type A and canonical form $N'$ \\
		$\Delta \vdash N \chk A [N']$ & Intro term N has type A and canonical form $N'$
	\end{tabular}
\end{center}

@@

\paragraph{Initialization}
$x \Leftarrow \kw{init} N; E$

\begin{mathpar}
	\inferrule
	{\Delta \vdash N \chk \type{Bool}
		\\ \Delta, x : \type{Qbit}; P \circ (x \mapsto \kw{state}(N)) \vdash E \syn y:B . Q}
	{\Delta; P \vdash x \Leftarrow \kw{init} N; E \syn y:B. (\exists x:\type{Qbit}.Q)}
\end{mathpar}
where x is fresh and \texttt{state} is a function that translates a classical representation of quantum state (such as boolean here) to the equivalent quantum state vector.

Initialization can only be performed for a term $N$ that can be type checked as a \type{Bool}. Then, we look at the rest of the computation, $E$, in a context that adds $x$ of type \type{Qbit} under a precondition that extends the previous precondition, $P$, with the strongest postcondition for initialization, that is, the newly bound variable $x$ points to the representation of a new qubit state; this should synthesize the strongest postcondition (with respect to the expanded precondition in the context), $Q$, for $E$.

@@

\paragraph{Measurement}
$x \Leftarrow \kw{measure} N; E$

\begin{mathpar}
	\inferrule
	{\Delta \vdash N \chk \type{Qbit}
		\\ \Delta; \Psi; P \prove (N \hookrightarrow -)
		\\ \Delta, x:\type{Bool}; P \circ ((N \mapsto -) \multimap \kw{emp}) \vdash E \syn y:B . Q}
	{\Delta; P \vdash x \Leftarrow \kw{measure} N; E \syn y:B. (\exists x:\type{Bool}.Q)}
\end{mathpar}

The measurement rule can be read in a similar way as the previous rule. But in measurement, we need to additionally prove that the verification condition (written as a sequent), the precondition $P$ implies that the qubit $N$ is allocated, holds in the heap context $\Psi$. The strongest postcondition here is that the fragment of heap that only refers to location N becomes empty.
